---
title: 'PH C240B: HW 2'
author: "Yue You"
output: 
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(survival)
library(dplyr)
library(doParallel)
library(mvtnorm)
```

## Problem 1

Consider the following data generating system for $(T,C)$, a random person’s survival time and censoring time in days. $T$ follows an exponential distribution with $\lambda = 0.01$ and $C$ follows a weibull distribution with shape and scale parameters, 4 and 100. Simulate 1000 random draws of $n = 1000$ from this distribution and for each draw compute the Kaplan-Meier estimator for survival curve at times 50, 60 and 70, where we consider the study ending at time 100. Form 95\% simultaneous confidence bands from each draw and check simultaneous coverage of the true survival. Use the influence curve for the KM estimator to obtain the CI’s for each draw. Report the coverage percentage. Note, use foreach for this as shown in lab3sol.Rnw on bcourses to save some time.

\vspace{.1in}

```{r, cache=TRUE}

S0 = function(t) 1-pexp(t,.01)

sim.cov = function(n) {
  T = rexp(n, .01)
  C = rweibull(n, 4, scale = 100)
  
  # record the observed right censored data
  Ttilde = pmin(T,C)
  Delta = T <= C & T < 100
  
  # do a KM fit
  km = survfit(Surv(Ttilde, Delta, type = "right") ~ 1, type = "kaplan-meier", 
               conf.int = .95)
  
  # make step functions
  km_step <- stepfun(km$time, c(1, km$surv))
  km_lower <- stepfun(km$time, c(1, km$lower))
  km_upper <- stepfun(km$time, c(1, km$upper))
  
  # order the times and corresponding deltas
  T_ord = Ttilde[order(Ttilde)]
  D_ord = Delta[order(Ttilde)]
  
  
  ##### CHANGES MADE ######
  # Get all known death times in order and compute the hazard at such times
  # lambda = km$n.event[D_ord]/km$n.risk[D_ord]
  T_death = T_ord[D_ord]
  nrisk = vapply(T_death, FUN = function(x) sum(Ttilde >= x), FUN.VALUE = 1)
  lambda = 1/nrisk
  #########################
  
  # compute the prob of surviving up to or past each death time
  Pbar = vapply(T_death, FUN = function(time) mean(Ttilde >= time), FUN.VALUE = 1)
  
  IC_time = function(Ttilde, Delta, time, n) {
    sum((-(T_death <= time)*km_step(time)/(1 - lambda))*
          lambda^2*(1 - lambda)^2/((1 - lambda)^2/n + lambda^2*Pbar)*
          ((Ttilde == T_death & Delta == 1)/lambda - (Ttilde > T_death)/(1 - lambda)))
  }
  
  # compute the IC matrix with each col an IC        
  IC = vapply(c(50,60,70),FUN = function(t) {
    unlist(lapply(1:n, FUN = function(i) {
      IC_time(Ttilde[i], Delta[i], t, n)
    }))
  }, FUN.VALUE = rep(1,n))
  
  
  # COMPUTE THE CORRELATION and draw the random three-d normals
  Sigma = cor(IC)
  z = rmvnorm(1e6, c(0,0,0), Sigma)
  
  # compute the max abs val of each of the 3-d normals then choose the 
  # 95th quantile of that vector, which is the simultaneous number of SE's
  z_abs = apply(z, 1, FUN = function(row) max(abs(row)))
  SE_num = quantile(z_abs, .95)
  
  # Note how the CI is wider when demanding simultaneous coverage
  Cov50 = km_lower(50) <= S0(50) & km_upper(50) >= S0(50)
  Cov60 = km_lower(60) <= S0(60) & km_upper(60) >= S0(60)
  Cov70 = km_lower(70) <= S0(70) & km_upper(70) >= S0(70)
  
  indy_cov = all(Cov50, Cov60, Cov70)
  
  # simultaneous for 40, 50, 60
  Sim50 = km_step(50) - SE_num*sd(IC[,1])*sqrt(n-1)/n <= S0(50) & 
    S0(50) <= km_step(50) + SE_num*sd(IC[,1])*sqrt(n-1)/n
  Sim60 = km_step(60) - SE_num*sd(IC[,2])*sqrt(n-1)/n <= S0(60) & 
    S0(60) <= km_step(60) + SE_num*sd(IC[,2])*sqrt(n-1)/n
  Sim70 = km_step(70) - SE_num*sd(IC[,3])*sqrt(n-1)/n <= S0(70) & 
    S0(70) <= km_step(70) + SE_num*sd(IC[,3])*sqrt(n-1)/n
  
  sim_cov = all(Sim50, Sim60, Sim70)
  
  return(c(indy_cov = indy_cov, sim_cov = sim_cov))
}

sim.cov(1000)

registerDoParallel(cores = detectCores())
getDoParWorkers()

B = 1000
n = 1000
ALL = foreach(i=1:B,.packages=c("mvtnorm"), .errorhandling = "remove") %dopar% {sim.cov(n)}
res = do.call(rbind, ALL)
colMeans(res)
```

Individual 95\% confidence intervals (indy_cov) only cover the truth simultaneously at lower than 95\% of the time. And simutaneous 95\% confidence intervals (sim_cov) covers roughly 95 \%. Simutaneous 95\% confidence intervals are also wider than individua confidence intervals because they need to cover a set of points opposed to a singular point. 
