---
title: 'PH C240B: Assignment 3'
author: "Rachael Phillips"
date: "11/9/2017"
output: 
  pdf_document: default
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(survival)
library(dplyr)
library(doParallel)
library(mvtnorm)
```

### Problem 1
Show the CAR condition, $x \rightarrow Pr(O = o | X = x)$ for $x \in \mathcal{C}(o)$ is constant implies $Pr(X=x| O = o) = Pr(X = x | x \in \mathcal{C}(o))$. You may assume all random variables here are discrete for simplicity.

\vspace{.1in}
Intuitively, the CAR assumption says that the observation of $O=o$ is not influenced by the specific value of $X$ in $\mathcal{C}(O)$ which was taken, only by the fact that $X$ did take a value in $\mathcal{C}(O)$. Thus, the CAR condition is equivalent to 

$$Pr(O = o | X = x) = Pr(O=o | X \in C(O)) \text{ for all }o, x\in\mathcal{C}(o)$$
which can be rewritten as, 

$$Pr(O = o | X = x \text{ and } X \in C(O)) = Pr(O=o | X \in C(O)) \text{ for all } x\in\mathcal{C}(o)\text{ .}$$

clearly identifying the conditional independence assumption: given $X\in\mathcal{C}(O)$, the events $X=x$ and $O=o$ are conditionally independent. By symmetry of conditional independence we have, 

$$Pr(X=x|O=o \text{ and } X \in \mathcal{C}(O)) = Pr(X=x|X\in\mathcal{C}(O))\text{ .}$$

But, since the former is equal to $Pr(X=x|O=o)$, we have that CAR is equivalent to 

$$ Pr(X=x|O=o) = Pr(X=x|X\in\mathcal{C}(O)) \text{ for all } x\in\mathcal{C}(o)\text{ .}$$

So we see that this conditional probability of the full data, $X$, given $O$ is only dependent the distribution of $P_X$
Thus the observation of $O=o$ tells us no more, in the sense of what is now the conditional distribution of $X$, than the obvious $X\in\mathcal{C}(O)$

### Problem 2 
Let $P_{X,\epsilon}$ be a path through $P_X$, the distribution of the full data, $X$, and having score $S_1(X)$. This then defines a path $P_{P_{X,\epsilon},G}$ through the observed data distribution, $P_{P_XG}$. Show that the scores generated by these paths are $E[S_1(X) | O = o]$.

\vspace{.1in}

$$\frac{d}{d\epsilon}logdP_{P_{X,\epsilon}G}/dP_{P_{X}G}|_{\epsilon=0} = \frac{d}{d\epsilon}logdP_{P_{X,\epsilon}G}|_{\epsilon=0}$$

$$ = \int\frac{dP(O=o|X=x)S_1(x)dP(x)d\nu(x)}{dP_{P_{X}G}(o)}$$

$$ = \int S_1(x)dP(x|O=o)d\nu(x)$$

$$ = E\big[S_1(X)|O=o\big] \text{ .}$$

### Problem 3
Let $G_{\epsilon}$ be a path through $G$, the distribution of the censoring time, $C$, given $X$, having score $S_2(C,X)$. This then defines a path $P_{P_XG_{\epsilon}}$ through the observed data distribution, $P_{P_XG}$. Show that the scores generated by these paths are $E[S_2(C, X) | O = o]$.

\vspace{.1in}

$$\frac{d}{d\epsilon}logdP_{P_XG_{\epsilon}}/dP_{P_{X}G}|_{\epsilon=0} = \frac{d}{d\epsilon}logdP_{P_XG_{\epsilon}}|_{\epsilon=0}$$
$$ = \int\frac{S_2(c,x)dP(O=o|X=x)dP(x)d\nu(x)}{dP_{P_XG_{\epsilon}}(o)}$$
$$ = \int S_2(c,x)dP(x|O=o)d\nu(x)$$
$$ = E\big[S_2(C,X)|O=o\big]\text{ .}$$

### Problem 4
This problem involves simulating data under a general Cox model. Let’s make the assumption we have a conditional hazard of death at time, $t$, given by $\lambda(t | X) =  \lambda_0(t)exp(f_{\beta}(X))$ where $X$ is a set of covariates and $f_{\beta}$ is a function indexed by $\beta$, say finite dimensional. Assume the baseline hazard is  $\lambda_0(t) = exp(rt)$ for positive $r$. Given $X$, what is the distribution of death times? Prove your answer.

\vspace{.1in}

```{r}

```

### Problem 5
Complete the first problem from LabCox in the lab section of the files on bCourses.

\vspace{.1in}

```{r}

```

### Bonus
Assume a CAR model for full data consisting of survival time, censoring time, the continuous baseline covariates and randomly assigned treatment indicator. We have observed data $min(T,C),\Delta$ along with the covariates and treatment indicator. Someome receives a data set of 1000 independent subjects drawn from this model from an RCT and runs a Cox Proportional hazards regression with treatment as the only covariate, showing a significantly negative coefficient. Can you convince this person he may be wrong via simulation? Explain how you set up your simulation and turn in your code to show the results.

\vspace{.1in}

```{r}

```

